Plug-and-Play Dual-Tree Algorithm Runtime Analysis

### 2015-12-28 bounds.tex
Add editor.
Add acknowledgements.
Update headings and abstract formatting.
Use titlecase for section headings.

\title{Plug-and-play dual-tree algorithm runtime analysis}

@@

\editor{Unknown}

@@

\begin{abstract}
%A large collection of machine learning algorithms contain subproblems that can
%be efficiently solved with trees.  In the situation where a dual-tree algorithm
%with cover trees is used to solve a problem, rigorous worst-case runtime bounds
%can often be shown.  A worst-case runtime bound is proven for all cover tree
%dual-tree algorithms; this bound can be employed to give useful runtime bounds
%for individual problems.  We show a new, tighter bound of $O(c^9 \nu N)$ for
%all-nearest-neighbor search and also provide linear runtime bounds for
%approximate kernel density estimation and exact range count/search under some
%mild assumptions.  These are the first linear runtime bounds we are aware of for
%range count/search.
%Numerous machine learning algorithms contain pairwise statistical problems at
%their core. Often, tree structures are used to solve these problems efficiently.
%Dual-tree algorithms using cover trees have rigorous worst-case runtime
%guarantees for some of these problems. In this paper, we present a {\em
%problem-independent} runtime guarantee for {\em any} dual-tree algorithm using
%the cover tree, separating out the problem-dependent and the problem-independent
%elements. This allows us to just plug in bounds for the problem-dependent
%elements to get runtime guarantees for dual-tree algorithms for any pairwise
%statistical problem. We demonstrate this plug-and-play procedure for
%nearest-neighbor search and approximate kernel density estimation to get
%improved runtime guarantees.  Under some assumptions, we also present the first
%linear runtime guarantee for dual-tree based range search.
Numerous machine learning algorithms contain pairwise statistical problems at
their core---that is, tasks that require computations over all pairs of input
points if implemented naively.  Often, tree structures are used to solve these
problems efficiently. Dual-tree algorithms can efficiently solve or approximate
many of these problems.  Using cover trees, rigorous worst-case runtime
guarantees have been proven for some of these algorithms. In this paper, we
present a {\em problem-independent} runtime guarantee for {\em any} dual-tree
algorithm using the cover tree, separating out the problem-dependent and the
problem-independent elements.  This allows us to just plug in bounds for the
problem-dependent elements to get runtime guarantees for dual-tree algorithms
for any pairwise statistical problem without re-deriving the entire proof. We
demonstrate this plug-and-play procedure for nearest-neighbor search and
approximate kernel density estimation to get improved runtime guarantees.  Under
mild assumptions, we also present the first linear runtime guarantee for
dual-tree based range search.
\end{abstract}

\section{Dual-tree algorithms}

@@

\section{Tree imbalance}
\label{sec:imbalance}

@@

\section{General runtime bound}
\label{sec:bound}

@@

\section{Nearest neighbor search}
\label{sec:nns}

@@

\section{Approximate kernel density estimation}
\label{sec:akde}

@@

\section{Range search and range count}
\label{sec:rs}

@@

>>>>

\title{Plug-and-Play Dual-Tree Algorithm Runtime Analysis}

@@

\editor{Nando de Freitas}

\jmlrheading{12}{2015}{1-29}{1/15}{12/15}{R.R. Curtin, D. Lee, W.B. March, P. Ram}

\ShortHeadings{Plug-and-Play Dual-Tree Algorithm Runtime Analysis}{Curtin, Lee, March,
and Ram}

@@

\begin{abstract}%
%A large collection of machine learning algorithms contain subproblems that can
%be efficiently solved with trees.  In the situation where a dual-tree algorithm
%with cover trees is used to solve a problem, rigorous worst-case runtime bounds
%can often be shown.  A worst-case runtime bound is proven for all cover tree
%dual-tree algorithms; this bound can be employed to give useful runtime bounds
%for individual problems.  We show a new, tighter bound of $O(c^9 \nu N)$ for
%all-nearest-neighbor search and also provide linear runtime bounds for
%approximate kernel density estimation and exact range count/search under some
%mild assumptions.  These are the first linear runtime bounds we are aware of for
%range count/search.
%Numerous machine learning algorithms contain pairwise statistical problems at
%their core. Often, tree structures are used to solve these problems efficiently.
%Dual-tree algorithms using cover trees have rigorous worst-case runtime
%guarantees for some of these problems. In this paper, we present a {\em
%problem-independent} runtime guarantee for {\em any} dual-tree algorithm using
%the cover tree, separating out the problem-dependent and the problem-independent
%elements. This allows us to just plug in bounds for the problem-dependent
%elements to get runtime guarantees for dual-tree algorithms for any pairwise
%statistical problem. We demonstrate this plug-and-play procedure for
%nearest-neighbor search and approximate kernel density estimation to get
%improved runtime guarantees.  Under some assumptions, we also present the first
%linear runtime guarantee for dual-tree based range search.
Numerous machine learning algorithms contain pairwise statistical problems at
their core---that is, tasks that require computations over all pairs of input
points if implemented naively.  Often, tree structures are used to solve these
problems efficiently. Dual-tree algorithms can efficiently solve or approximate
many of these problems.  Using cover trees, rigorous worst-case runtime
guarantees have been proven for some of these algorithms. In this paper, we
present a {\em problem-independent} runtime guarantee for {\em any} dual-tree
algorithm using the cover tree, separating out the problem-dependent and the
problem-independent elements.  This allows us to just plug in bounds for the
problem-dependent elements to get runtime guarantees for dual-tree algorithms
for any pairwise statistical problem without re-deriving the entire proof. We
demonstrate this plug-and-play procedure for nearest-neighbor search and
approximate kernel density estimation to get improved runtime guarantees.  Under
mild assumptions, we also present the first linear runtime guarantee for
dual-tree based range search.
\end{abstract}

\section{Dual-tree Algorithms}

@@

\section{Tree Imbalance}
\label{sec:imbalance}

@@

\section{General Runtime Bound}
\label{sec:bound}

@@

\section{Nearest Neighbor Search}
\label{sec:nns}

@@

\section{Approximate Kernel Density Estimation}
\label{sec:akde}

@@

\section{Range Search and Range Count}
\label{sec:rs}

@@

\section*{Acknowledgements}

The authors gratefully acknowledge the helpful and insightful comments of the
anonymous reviewers.
